\section{Process}
As it can be seen in Chapter \ref{cha:tasks}, we set up a series of tasks in order to reach our overall goal and thereby accomodate our problem statement. These task have a linear dependency which have led us to follow a waterfall inspired work process. This process is seperated into two major parts; Design and implementation of our framework, and comparison between our framework and other frameworks. 

During design and implementation, some features might be iterated multiple times, but when this happends it will be made clear. Before we start the comparison part, we will have completed the design and implementation part.

The parts each cover multiple sub parts, chosen to perform the tasks mentioned in Chapter \ref{cha:tasks}. The report follows the structure of these parts to document the process.\todo{gør dette sandt.} The parts and sub parts consist of:
\begin{description}
\item[Design and implementation of our framework] \hfill
\begin{itemize} 
\item Research of related works.
\item Designing our framework.
\item Implementation of our framework.
\item Test of our framework.
\end{itemize}
\item[Comparison between our framework and other frameworks] \hfill
\begin{itemize} 
\item Research of comparison methology.
\item Implementation of demo application in our framework.
\item Implementation of the same demo applications in other frameworks.
\item Comparing implementations.
\item Comparing the frameworks.
\item Evaluation of our framework.
\end{itemize}
\end{description}
