\section{SYCL}
\textit{SYCL} is a high-level programming language that provide an abstraction layer for \textit{OpenCL} and it is developed by \textit{Khronos Group}. This section describes \textit{SYCL} based on the information available at the \textit{Khronos Group} website \cite{khronosSYCL}.

\subsection{Goals}
As opposed to regular \textit{OpenCL} development, \textit{SYCL} enables the host and device code to be contained within a single source. \textit{SYCL} exposes the \textit{OpenCL} feature-set with a higher abstraction level, as well as most modern \textit{C++} features. 

The \textit{Khronos Group} aim to follow the current \textit{C++} standard developments and integrate it with \textit{OpenCL} features.

\subsection{Programming model}
\textit{OpenCL} has commands for memory object creation, copying, mapping and synchronisation. \textit{SYCL} wraps these as a \textit{command group} that can manage these commands. Listing \ref{code:saxpySycl} shows an implementation of the \textit{SAXPY} computation. \textit{SYCL} need to know which variables should be available for device use, and this is indicated by firstly setting up host storage as seen through line \ref{code:saxpySycl:setupS} to \ref{code:saxpySycl:setupE}. These variables, \texttt{x}, \texttt{y}, and \texttt{z}, are then placed in a buffer, as seen on lines \ref{code:saxpySycl:buffersS} to \ref{code:saxpySycl:buffersE}, which marks the data to be shared between host and device and initializes the queue. Next, at lines \ref{code:saxpySycl:selector} and \ref{code:saxpySycl:queue}, the available devices are registered by initializing the \texttt{device\_selector}. At line \ref{code:saxpySycl:queueSubmit} the buffer elements are submitted by the \texttt{cgh} handler, and at the following three lines it is specified how each buffer element should be accessed; \texttt{x} and \texttt{y} is set with the \textit{read} access mode and \texttt{z} is set with the \textit{discard\_write} access mode. The actual execution of \textit{SAXPY} is specified at lines \ref{code:saxpySycl:exec1} and \ref{code:saxpySycl:exec2}.

\begin{lstlisting}[caption={\textit{SAXPY} implemented in \textit{SYCL}.}, label={code:saxpySycl}]
sycl::float4 x = { 1.0, 2.0, 3.0, 4.0 }; ~\label{code:saxpySycl:setupS}~
sycl::float4 y = { 2.0, 3.0, 4.0, 5.0 };
sycl::float4 z = { 0.0, 0.0, 0.0, 0.0 }; ~\label{code:saxpySycl:setupE}~

sycl::buffer<sycl::float4, 1> a_sycl(&x, sycl::range<1>(1)); ~\label{code:saxpySycl:buffersS}~
sycl::buffer<sycl::float4, 1> a_sycl(&y, sycl::range<1>(1));
sycl::buffer<sycl::float4, 1> a_sycl(&z, sycl::range<1>(1)); ~\label{code:saxpySycl:buffersE}~

sycl::default_selector device_selector; ~\label{code:saxpySycl:selector}~
sycl::queue queue(device_selector); ~\label{code:saxpySycl:queue}~

queue.submit([&] (sycl::handler& cgh) { ~\label{code:saxpySycl:queueSubmit}~
  auto x_acc = x_sycl.get_access<sycl::access::mode::read>(cgh);
  auto y_acc = y_sycl.get_access<sycl::access::mode::read>(cgh);
  auto z_acc = z_sycl.get_access<sycl::access::mode::discard_write>(cgh);

  cgh.single_task<class saxpy>([=] () { ~\label{code:saxpySycl:exec1}~
    z_acc[0] = 2 * x_acc[0] + y_acc[0]; ~\label{code:saxpySycl:exec2}~
  });
});
\end{lstlisting}

Even though \textit{SYCL} provides a higher abstraction level compared to regular \textit{OpenCL}, low-level \textit{C++} and \textit{OpenCL} features are still available.

\subsection{Implementation}
\textit{SYCL} primarily targets \textit{OpenCL}, but can also target other backends as it is translated to the \textit{SPIR-V} intermediate representation.

\subsection{Key Points}
A key point of \textit{SYCL} is its way of specifying data that should be available for host and devices. By placing data in the buffers and specifying how the data should be accessed, shown in listing \ref{code:saxpySycl} as \texttt{read} and \texttt{discard\_write}, the availability of data is clear. 
