\section{Queuing Actions} \label{cha:queueingActions}
To modify a \texttt{yagal::Vector} we create a concept of actions. These actions contain the necessary information for generating a kernel, that executes the needed logic.

An action is an intended task to be performed on the \texttt{yagal::Vector}, and is constructed as a result of a developer calling a function on the \texttt{yagal::Vector}. For instance, calling \texttt{myVector.add(5)} will result in an \texttt{AddAction} being created with a copy of the parameters, 5 in this example, needed to perform the calculation. The developer must then call \texttt{myVector.exec()} to perform all actions. We use the actions as the source of our code generation, as we generate code to perform exactly the content of the action.

The template class \texttt{Vector<T>} is expanded, as seen in Listing \ref{code:actionAdditions}. A vector of \texttt{Action<T>} elements are added to the fields at line \ref{code:actionAdditions:vec}, it contains the actions needing to be performed on the \texttt{Vector<T>}, in the order they were added. Functions have been added to generate actions that are placed in this vector. An example of this is \texttt{Vector<T>\& add(T value)} at line \ref{code:actionAdditions:val}, which creates an action that represents adding a value to all elements in the \texttt{Vector<T>}, and places it at the back of the \texttt{\_actions} vector. The other added function is \texttt{Vector<T>\& exec()}, which is the function that consumes the action vector, generates \textit{LLVM IR} based on the consumed actions, translates the \textit{LLVM IR} to \textit{PTX}, loads the \textit{PTX} on the device, and executes it. The content of the \texttt{Vector<T>\& exec()} function is not the focus of this section, and is explained in section \ref{cha:puttingStepsTogether}.

\begin{lstlisting}[caption={Vector<T> action additions}, label={code:actionAdditions}]
namespace yagal{
    template <typename T>
    class Vector{
    private:
        /* omitted fields */

        std::vector<std::shared_ptr<internal::Action>> _actions; ~\label{code:actionAdditions:vec}~

        /* omitted functions */
    
    public:
        Vector<T>& add(T value) {~\label{code:actionAdditions:val}~
            _actions.emplace_back(new internal::AddAction<T>(value));
            return *this;
        }
        
        /* omitted functions */

        Vector<T>& exec(){
            /* omitted logic */
        }
    }
}
\end{lstlisting}

The template class \texttt{Action<T>} is the single top level class in a hierarchy, where every bottom level class is a concrete action that can be performed on a \texttt{Vector<T>}. These actions are grouped by mid level classes, that define their input parameters. An example is the \texttt{AddAction<T>} being a \texttt{SimpleAction<T>}, as it takes a single input value to perform the action on an element of the \texttt{Vector<T>}, and the \texttt{SimpleAction<T>} being an \texttt{Action<T>} to allow us to contain it with other actions in the  vector on line \ref{code:actionAdditions:vec}.

\todo{create some class diagram}
