\section*{Reading Instructions}
This thesis is meant to reflect and document the work of group dpw106f18 at their 10th semester at Aalborg University. This report have been written during spring 2018, and the material presented is based upon what was available at this time.

The reference material is available in the bibliography at the end of the report, and includes links to official documentation, websites, articles, scientific papers, and books. 

The intended audience for this thesis is readers with some previous experience or understanding of both GPGPU development and \textit{C++}.

A huge thanks to our supervisor, Lone Leth Thomsen from Aalborg University's department of computer science. Even tough we must have been a difficult group to work with, we appreciate all the help and support you provided throughout the thesis. Thank you!

This thesis is intended to be read in sequential order, as the chapters correspond to the tasks we performed in the order we worked on them. These chapters are;
\begin{description}
    \item[Introduction]\hfill\\
    This chapter contains the motivation and problem statement of the thesis, and in addition covers our development process and prerequisites for the thesis.
    \item[Related Works]\hfill\\
    This chapter contains our investigation of related works that are similar to ours.
    \item[Design Principles]\hfill\\
    This chapter contains the design guidelines that we followed to during development.
    \item[Framework Design]\hfill\\
    This chapter contains the design of our framework, including our approach, the design of the API, and the design of the underlying architecture.
    \item[Framework Implementation]\hfill\\
    This chapter contains documentation of our implementation and decisions taken during the implementation.
    \item[Problems] \hfill\\
    This chapter contains the biggest issues that was encountered during development, and our proposals for how these issues can be solved.
    \item[Framework Demo and Comparison]\hfill\\
    This chapter contains a presentation of our comparison methodology, how we used it, and an evaluation of our work and of the related works.
    \item[Reflection]\hfill\\
    This chapter contains our reflections upon the thesis as a whole, which includes our choice of related works, our design, our choice of technologies, our implementation, and our comparison.
    \item[Conclusion]\hfill\\
    This chapter contains the conclusion of the thesis.
    \item[Future work]\hfill\\
    This chapter contains a description of how this thesis could move forward, if development were to continue.
\end{description}